Ludwig User Tutorial

### 2019-08-14 docs/intro.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  intro.tex
%
%  Contains introduction and quick start for users.
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2016-2018 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

This section provides a brief overview of compiling and running
the code, which should allow users to get off the ground.
The section assumes the code has been downloaded and you are in the
top level directory.

\subsection{Configuration}

You will need to copy one of the existing configuration files from
the \texttt{config} directory to the top level directory. This file
sets all the relevant local details for compilation and so on. For
example:
\begin{lstlisting}
$ cp config/lunix-gcc-default.mk config.mk
\end{lstlisting}
Edit the new \texttt{config.mk} file to check, or adjust, the options
as required. You will need to identify your local C compiler. The
example uses \texttt{gcc}
in serial, and \texttt{mpicc} in parallel. For MPI in particular,
local details may vary. A quick way to compile the code is via
\begin{lstlisting}
$ cd tests
$ make compile-mpi-d3q19
\end{lstlisting}
This integrates a number of separate stages which are described
indivually in the following sections. The resulting executable
will be \texttt{./src/Ludwig.exe} at the top level.

\subsubsection{targetDP}

A \texttt{targetDP} layer is required by the code in all cases. It
allows compilation for both simple CPU systems and
for a number of accelerator systems using a single source code. This
will make the appropriate choices depending what has been specified
in the top-level \texttt{config.mk} file. From the top-level directory
\begin{lstlisting}
$ cd targetDP
$ make
\end{lstlisting}
should produce the appropriate \texttt{libtarget.a} library.

\subsubsection{Serial}
\label{section:quick-serial}

First, compile the MPI stub library in the \texttt{mpi\_s}
directory. You should be able to build the library and run the
tests using, e.g.:

\begin{lstlisting}
$ make libc
gcc -Wall -I.  -c mpi_serial.c
ar -cru libmpi.a mpi_serial.o
$ make testc
gcc -Wall   mpi_tests.c -L. -lmpi
./a.out
Running mpi_s tests...
Finished mpi_s tests ok.
\end{lstlisting}

Now compile the main code in the \texttt{src} directory.
Compilation should look like:

\begin{lstlisting}
$ make serial
make serial-d3q19
make serial-model "LB=-D_D3Q19_" "LBOBJ=d3q19.o"
make lib
make libar "INC=-I. -I../mpi_s" "LIBS= -L../mpi_s -lmpi -lm"
gcc -D_D3Q19_ -Wall -O2 -I. -I../mpi_s -c model.c
gcc -D_D3Q19_ -Wall -O2 -I. -I../mpi_s -c propagation.c
...
\end{lstlisting}
which shows that the default lattice Boltzmann model is D3Q19.
Successful compilation will provide an executable \texttt{Ludwig.exe}
which is linked against the MPI stub library.

\subsubsection{Parallel}

Here, you do not need to compile the stub library. Simply compile
the main code with, e.g.,:
\begin{lstlisting}
$ make mpi
make mpi-d3q19
make mpi-model "LB=-D_D3Q19_" "LBOBJ=d3q19.o"
make libmpi
make libar "CC=mpicc"
mpicc -D_D3Q19_ -Wall -O2 -I. -c model.c
mpicc -D_D3Q19_ -Wall -O2 -I. -c propagation.c
...
\end{lstlisting}

Again, this should produce an executable \texttt{./Ludwig.exe}
in the current directory, this time linked against the true MPI
library.

@@

Make a copy of the input file
\begin{lstlisting}
$ cp ../tests/regression/d3q19/serial-auto-c01.inp .
\end{lstlisting}

@@

So, the first few lines of the above file are:
\lstinputlisting[firstline=1, lastline= 16]
{../tests/regression/d3q19/serial-auto-c01.inp}
Here, the first key value pair \texttt{N\_cycles 40} sets the number
of time steps to 40, while the second, \texttt{size 64\_64\_64}, sets
the system size to be 64 points in each of the three coordinate
directions. The \texttt{grid} key relates to the parallel
domain decomposition as is discussed in the following
section.

@@

\begin{lstlisting}[belowskip=0pt]
$ ./Ludwig.exe ./serial-auto-c01.inp
\end{lstlisting}
\lstinputlisting[aboveskip=0pt, firstline=1, lastline= 16]
{../tests/regression/d3q19/serial-auto-c01.log}
This output shows that the appropriate input file has been read, and
the system size set correspondingly with a number of default settings
including periodic boundary conditions. ``No free energy'' tells us we
are using a single Newtonian fluid.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  intro.tex
%
%  Contains introduction and quick start for users.
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2016-2019 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

Make a copy of the input file
\begin{lstlisting}
$ cp ../tests/regression/d3q19-extra/serial-auto-c01.inp .
\end{lstlisting}

@@

So, the first few lines of the above file are:
\lstinputlisting[firstline=1, lastline= 16]
{../tests/regression/d3q19-extra/serial-auto-c01.inp}
Here, the first key value pair \texttt{N\_cycles 40} sets the number
of time steps to 40, while the second, \texttt{size 64\_64\_64}, sets
the system size to be 64 points in each of the three coordinate
directions. The \texttt{grid} key relates to the parallel
domain decomposition as is discussed in the following
section.

@@

\begin{lstlisting}[belowskip=0pt]
$ ./Ludwig.exe ./serial-auto-c01.inp
\end{lstlisting}
\lstinputlisting[aboveskip=0pt, firstline=1, lastline= 16]
{../tests/regression/d3q19-extra/serial-auto-c01.log}
This output shows that the appropriate input file has been read, and
the system size set correspondingly with a number of default settings
including periodic boundary conditions. ``No free energy'' tells us we
are using a single Newtonian fluid.

### 2019-08-14 docs/tutorial/tutorial.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  tutorial.tex
%
%  Introductory tutorial for new users of Ludwig. 
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre and
%  Department of Physics, University of Strathclyde
%
%  Contributing authors:
%  Oliver Henrich (oliver.henrich@strath.ac.uk)
%  Fraser Mackay (s1026487@sms.ed.ac.uk)
%
%  (c) 2008-2017 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{lstlisting}
$ ls ludwig
LICENSE Makefile.mk config docs mpi_s src target tests util version.h
\end{lstlisting}

@@

\begin{lstlisting}[style=terminalverbatim]
$ cd /ludwig/docs/
$ make pdf
\end{lstlisting}

@@

\begin{lstlisting}[style=terminalverbatim]
$ cd /ludwig/docs/tutorial
$ make
\end{lstlisting}

@@

\begin{enumerate}
\item Go to the top level directory \texttt{/ludwig}: \\
\begin{lstlisting}
$ cd ludwig
\end{lstlisting}
\item Copy the configuration file \texttt{lunix-gcc-default.mk} to top level directory 
and rename to \texttt{config.mk} (both steps exectued below at once): \\
\begin{lstlisting}
$ cp config/lunix-gcc-default.mk ./config.mk
\end{lstlisting}
Note that you may need to specify the path to the MPI wrapper compiler in this file.
\item Issue \texttt{make} in \texttt{/target}: \\
\begin{lstlisting}
$ cd target/
$ make 
\end{lstlisting}
This creates dummy functions for calls to the targetDP library.
\item Issue \texttt{make} in \texttt{/mpi\_s}: \\
\begin{lstlisting}
$ cd ../mpi_s/
$ make 
\end{lstlisting}
This creates dummy functions for calls to the MPI-library,
which is required for serial compilation.
\item Edit \texttt{Makefile} in \texttt{/src}. To switch off 
the assertions set \texttt{OPTS = -DNP\_D3Q6 -DNDEBUG}. This is
important for production runs (see also comment in \texttt{Makefile}): \\
\begin{lstlisting}
$ cd ../src/
$ vim Makefile
...

include ../Makefile.mk

MAIN = main
EXECUTABLE = Ludwig.exe
LIBRARY = libludwig.a

OPTS = -DNP_D3Q6 -DNDEBUG
LIBS = -L../target -ltarget -lm
INC = -I. -I ../target
...
\end{lstlisting}
\item To compile the code in serial issue: \\
\begin{lstlisting}
$ make serial
\end{lstlisting} 
To compile the parallel version of the code issue: \\
\begin{lstlisting}
$ make mpi
\end{lstlisting} 
This creates the executable file \texttt{Ludwig.exe} in \texttt{/src}. 
Note that more than one thread can be used in the compilation process,
which can lead to a considerable speedup  
\begin{lstlisting}
$ make -j4 mpi
\end{lstlisting} 
Before going from a serial to a parallel compilation or vice versa
you should clean the directory by issuing
\begin{lstlisting}
$ make clean
\end{lstlisting}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  tutorial.tex
%
%  Introductory tutorial for new users of Ludwig. 
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre and
%  Department of Physics, University of Strathclyde
%
%  Contributing authors:
%  Oliver Henrich (oliver.henrich@strath.ac.uk)
%  Fraser Mackay (s1026487@sms.ed.ac.uk)
%
%  (c) 2008-2019 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{lstlisting}
$ ls ludwig
CHANGES.md      Makefile        config          src             util
CONTRIBUTING.md Makefile.mk     docs            target          version.h
LICENSE         README.md       mpi_s           tests
\end{lstlisting}

@@

\begin{lstlisting}[style=terminalverbatim]
$ cd ./ludwig/docs/
$ make pdf
\end{lstlisting}

@@

\begin{lstlisting}[style=terminalverbatim]
$ cd ./ludwig/docs/tutorial
$ make
\end{lstlisting}

@@

\begin{enumerate}
\item Go to the top level directory \texttt{/ludwig}: \\
\begin{lstlisting}
$ cd ludwig
\end{lstlisting}
\item Copy the (parallel) configuration file \texttt{lunix-mpicc-default.mk}
to top level directory 
and rename to \texttt{config.mk} (both steps exectued below at once): \\
\begin{lstlisting}
$ cp config/unix-mpicc-default.mk ./config.mk
\end{lstlisting}
Note that you may need to specify the correct path to the MPI wrapper compiler
in this file (assumed to be \texttt{mpicc}).
\item Issue \texttt{make} in the top level directory: \\
\begin{lstlisting}
$ make 
\end{lstlisting}
This creates the executable file \texttt{./Ludwig.exe} in \texttt{/src}. 
Note that more than one thread can be used in the compilation process,
which can lead to a considerable speedup  
\begin{lstlisting}
$ make -j 4
\end{lstlisting} 
If you want a serial build, use the conifguation file
\texttt{config/unix-gcc-default.mk} and include the extra step
\begin{lstlisting}
$ make serial
$ make
\end{lstlisting}
